A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-11-27 01_Introduction.Rmd
necessary additions to the classical narrative

In a nutshell, the classic narrative of the development of Bronze Age burial rites in Europe presented e.g. by @hardingEuropeanSocietiesBronze2000a (following @hauslerBestattungssittenFruhenBronzezeit1977; @hauslerGrabUndBestattungssitten1994; @hauslerTotenorientierungUndGeographischer1996)  can be summarised as follows: In the early and middle Bronze Age inhumations dominated in different variations. Cremation burials in this time only occur in the Hungarian lowlands. Burial mounds are concentrated in parts of the Balkans as well as in eastern, western and northern Europe, while in central and southern Europe flat graves - mostly burials - predominate. In the Middle Bronze Age, barrows gained importance in Western and Central Europe -- a phenomenon highly linked to the emergence of the Tumulus culture in Central Europe. In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of Urnfield culture.

>>>>

In a nutshell, the classic narrative of the development of Bronze Age burial rites in Europe presented e.g. by @hardingEuropeanSocietiesBronze2000a (following @hauslerBestattungssittenFruhenBronzezeit1977; @hauslerGrabUndBestattungssitten1994; @hauslerTotenorientierungUndGeographischer1996)  can be summarised as follows: In the early and middle Bronze Age inhumations dominated in different variations. Cremation burials in this time only occur in the Hungarian lowlands. Burial mounds are concentrated in parts of the Balkans as well as in eastern, western and northern Europe, while in central and southern Europe flat graves - mostly burials - predominate. In the Middle Bronze Age, barrows gained importance in Western and Central Europe -- a phenomenon highly linked to the emergence of the Tumulus culture in Central Europe. In the late Bronze Age, the Urnfield period, cremation became the most common burial custom in the context of Urnfield culture. This presentation is, even when taking in account it's enormous simplification, partially misleading. Cremation was a lot more common already in the early and middle Bronze Age, though not in the context of the later Urnfield culture (e.g. in the Benelux area [@fokkensBronzeAgeLow2013] or Great-Britain [@robertsBritainIrelandBronze2013]).
